Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface

### 2022-02-04 doc/AMS-15/AMS-15_paper.Rmd
consolidate introduction

The gas-capable 134-sample MCSNICS is produced by NEC. Sample CO2 is introduced onto a titanium frit in an aluminum target using an arm with a probe that mates with the back of the target.

NEC hybrid gas ion source

NEC inlets

MICADAS source and inlets

Comparison of open split to closed method

Most interfaces to gas-accepting sputter ion sources are a closed inlet design. A mixture of sample and helium is introduced to the source through a capillary, and flow is controlled by the pressure in the sample volume and the length and diameter of the capillary. Flow may be adjusted by changing the pressure by varying the sample volume. We designed a simpler open inlet where a capillary samples gas in a flowing stream that exhausts to the atmosphere. This defines the delivery pressure as 1 ATM and therefore flow is controlled by the capillary and is not adjustable. The flow of CO2 to the source may be varied by controlling the ratio of CO2 to helium in the sample stream. An open inlet has the advantage of simplicity; flow of CO2 to the source is preset and there are no separate sample transfer and dilution steps. A disadvantage of the open split is that the flow of CO2 to the source varies with sample size (CO2:He) and changes over the course of a measurement due to dilution of the sample by displacement helium. Salazar and Szidat, Roberts gasbench, other open split methods.

Pros/cons

@@

The open split interface supplies a helium-CO2 mixture to a split where a capillary delivers gas to the source and the rest is vented. The current split is a VICI Luer adapter on the outflow side of the dual needle (Figure 1). The split has fairly low volume. Testing is needed to determine minimum sample flow to avoid atmosphere entering the split.

@@

Currently, 250cm of 50um capillary defines flow rate from the open split into the source, with a bulkhead connector at the source cage.

@@

Our method uses an open split and glass capillary to control gas flow to the ion source. Compared to closed inlet systems which regulate introduction of a CO2/Helium mixture to the ion source by regulating sample pressure, our method is simpler and faster, but does not maintain a constant flow of CO2 to the source. We compensate for this by limiting CO2 flow to a range that produces relatively constant C- ion current and stable ratios.

@@

Reference: gilson \@ref(Roberts 2013)

Carbonate samples are weighed into 7 ml septa-sealed Vacutainer vials (Labco \# 768W). Sealed vials are evacuated, then 1 ml 85% H3PO4 is added. The samples are allowed to react at 60C for 4-6 h. Sample pressure was adjusted to 1 ATM by adding helium or removing sample using a syringe.

Process blank samples were prepared from powdered C-1 or TIRI-F. Standards were prepared on the Gilson prep system by evacuating vials and back filling with 1 ATM of a modern standard gas.

>>>>

The gas-capable 134-sample MCSNICS from National Electrostatics Corporation (NEC) works on the same principal as hybrid sources developed by ETH (REF) and HVEE (REF). Sample CO2 is introduced onto a titanium frit in an aluminum target using stainless tubing mounted to an arm with a probe that mates with the back of the target. A CO~2~ - helium mixture is exposed to a beam of charged cesium atoms, which causes dissociation of CO~2~ and ionization of molecular and atomic species. Negative ions are extracted from the target and a magnetic filter picks species with masses corresponding to carbon.

Most interfaces to gas-accepting sputter ion sources are a closed inlet design. A mixture of sample and helium is introduced to the source through a capillary, and flow is controlled by the pressure in the sample volume and the length and diameter of the capillary. Flow may be adjusted by changing the pressure by varying the sample volume.

We designed a open inlet where a capillary samples gas for delivery to the ion source in a flowing stream that exhausts to the atmosphere. This defines the delivery pressure as 1 ATM and therefore flow is controlled by the capillary and is not adjustable. The flow of CO2 to the source may be varied by controlling the ratio of CO2 to helium in the sample stream. An open inlet has the advantage of simplicity; flow of CO2 to the source is preset and there are no separate sample transfer and dilution steps. This may translate to greater throughput by eliminating the time required for sample transfer, dilution, and setting of delivery pressure. A disadvantage of the open split is that the flow of CO2 to the source varies with sample size (CO2:He) and changes over the course of a measurement due to dilution of the sample by displacement helium. Our method compensates for this by limiting CO2 flow to a range that produces relatively constant C- ion current and stable ratios. Salazar and Szidat, Roberts gasbench, other open split methods.

@@

The open split interface supplies a helium-CO2 mixture to a split where a capillary delivers gas to the source and the rest is vented. A flow of helium is used to displace sample CO2 to the delivery capillary in the split. The core of the design is a dual needle and split housed in a VICI cross. Displacement helium flows in through one side of the cross, into the sample vial through 1/16" stainless steel tubing, which forms the outer flow path of the concentric dual needle. The mixture of sample CO2 and displacement helium flows up a needle inserted through the cross and stainless tubing which forms the inner flow path of the dual needle. The split is made from a VICI Luer adapter fitted to the outflow of the dual needle with the source delivery capillary inserted(Figure 1). The split has fairly low volume (< 1 mL). Testing is needed to determine minimum sample flow to avoid atmosphere entering the split. TODO: move flow ratio test to results.

@@

250cm of 50um capillary defines flow rate from the open split into the source, with a bulkhead connector at the source cage.

@@

@@

Reference: gilson

Carbonate samples are weighed into 7 ml septa-sealed Vacutainer vials (Labco \# 768W). Sealed vials are evacuated, then 1 ml 85% H3PO4 is added. The samples are allowed to react at 60C for 4-6 h. Sample pressure is adjusted to 1 ATM by adding helium or removing sample using a syringe. Evacuation, acid addition, and heated reaction are performed by the automated gas handling system developed for the CFAMS GIS carbonate system \@ref(Roberts 2013).

Process blank samples were prepared from powdered C-1 or TIRI-F. Standards were prepared using the automated gas handling system by evacuating vials prior to filling with 1 ATM of a modern standard gas.
